Comparing vector-based and ACT-R memory models using large-scale datasets:\protect\\User-customized hashtag and tag prediction on Twitter and StackOverflow

### 2014-09-21 writeups/dissertation/dissertation.tex
foo

% Default apa6 formatting for a CI
\newcommand{\CI}[2]{95\% CI [#1, #2]}

@@

These two memory systems were compared across two domains where users generate tags for posts: Twitter and StackOverflow.
Twitter is a microblogging service where users create 140 character tweets and broadcast those messages out to the users who follow their posts.
A user is free to annotate important concepts or words in the tweet by preceding words with hashtags ``\#''.
After the tweet is created the hashtags become hyperlinks, and if clicked on will take a user to a summary feed of all of the tweets across Twitter that have used that hashtag.
So these user-created hashtags on Twitter help connect the tweet to other tweets about the same topic through common hashtag use.
Some recent and often used hashtags for Twitter are \emph{\#photography}, \emph{\#startup}, \emph{\#4change}, \emph{\#android}, and \emph{\#solar}.

@@

Another issue with LSA is that the SVD matrix decomposition technique is computationally very expensive.
As the size of the original word by document matrix increases, the SVD computation becomes more and more time consuming.
If the domains where LSA is applied contained small enough datasets where SVD computation could be computed in a reasonable amount of time, then this would not be much of a concern.
However, with large human-created datasets such as Twitter, StackOverflow, and Wikipedia, the total size of the dataset far exceeds the maximum size that can even be computed by an LSA approach.
\textcite{Budiu2007} used the first six million pages of the Stanford WebBase corpus to compare performance between LSA and PMI on the TOEFL, and the LSA model could not be implemented with this size training dataset.
Instead, they compared the measures on the TOEFL by training on the smaller Touchstone Applied Science Associates (TASA) corpus created from \num{60527} samples of text from high school to early college textbooks.
This is one of the available corpora on the LSA website and is aimed to summarize the general reading up to the 1st year of college \parencite{Budiu2007}.

@@

With vector-based systems, order information is included by creating an environment vector ($e_{bind(i,\Phi)}$) that is a function of environment vectors for both the word ($e_{i}$) and location ($\Phi$).
This function has the useful property that the resulting environment vector and the location vector can be inverted to return the original word's environment vector.
Circular convolution and inverse circular convolution (i.e., circular correlation) are used by \textcite{Plate1995} and \textcite{Jones2007} to implement these operations.
This allows for retrieval requests such as ``playing \#$\Phi$'', ``just read \#$\Phi$'', and ``vector-based $\Phi$ systems''.

@@

Using decoding, the word-order memory vector for martin is deconvolved to the right ($o_{martin} \circledast \Phi$) which returns the environment vector for luther ($e_{luther}$).
Using resonance, a memory vector for ``martin $\Phi$'' is constructed ($e_{martin} * \Phi$), which correlates highest with the word-order memory vector for luther ($o_{luther}$).
Either way, luther is returned as the most likely term to match the phrase ``martin $\Phi$''.

@@

For example, suppose that the words ``stack overflow'' appear in a sentence, and I want to represent that ``stack'' preceded the word ``overflow''.
The ``stack'' environment vector ($e_{stack}$) is shifted one to the left, producing an environment vector for ``stack'' preceding a word ($e_{stack^{-1}}$).
That environment vector is then added to the memory vector for ``overflow'' ($m_{overflow} = m_{overflow'} + e_{stack^{-1}}$), so that ``overflow'' remembers that ``stack'' has preceded it in a sentence.

@@

The fan effect describes a behavior of human memory where memory items are more difficult to retrieve if they have occurred in a larger number of different contexts.
\textcite{Rutledge2008} used \textcite{Anderson1974} classic experimental paradigm to test the fan effect.
In the task, participants are exposed to a set of person and place pairs (e.g., ``the hippie is in the park'').
They are later asked to recall if a particular person and place pair was presented to them before.
The main manipulation is the ``fan'' of the person and place for each pair.
The fan is defined as the number of different contexts that each item in the pair is presented in.
For example, if items were presented to the participant where the ``hippie'' was both in the ``park'' and in the ``store'', then ``hippie'' would have a fan of two.
Recall accuracy decrease and time to recall increase as total fan of the person and pair increases.

@@

This preliminary research suggests that model fit can improve when tag predictions are customized to the specific user's past tagging history.
It also seems plausible that model accuracy can improve an even greater amount by using the time-dependent form of the base-level component in Table \ref{tabACTRModel},
in order to account for the fact that a user's tagging preferences change over time.

@@

A additional popular-hashtags dataset was collected for Twitter.
The purpose of this dataset is to test the accuracy of each model's predicted most likely hashtag(s) for a tweet.
The dataset contains approximately three million tweets where at least one of the 400 most popular hashtags are used, and was collected in real time until the tweet quota was reached.

@@

``Stop words'' (e.g., ``the'', ``a'', ``or'') are commonly used words that co-occur with all tags (i.e., have little to no discriminating power).
It is typically recommended \parencite{Bird2009} to remove stop words when building the co-occurrence matrix.
Otherwise the matrix may not scale as well since these low-predictor stop words commonly appear in the text.
Also for vector-based systems, memory vectors will become overly saturated with the environment vectors of the commonly-occurring stop words.

@@

For both equations, $i$ represents each chunk (e.g., a word), $j$ is a specific observation for a chunk (e.g., the 2nd time that ``a'' appears), and $n$ is the number of times each chunk has been observed.
Since the optimized learning form assumes a constant presentation rate for each chunk,
it only has to keep track of the time since the first presentation of a chunk ($L$) and the total number of instances observed for that chunk ($n$). 
It does not have to keep track of the time each specific chunk instance was observed.
Consequently, the optimized learning form of $B_{i}$ is computationally much less expensive than the standard form, and is actually set to the default in the ACT-R Common Lisp implementation.

@@

The base-level ($B_{i}$) component in Table \ref{tabACTRBLLModel} is different than the $B_{i}$ used in \textcite{Stanley2013} (Table \ref{tabModACTRModel}).
\citeauthor{Stanley2013} assumed no time dependence, no decay rate, and that the $B_{i}$ is simply a log-odds computation of the probability that each chunk is likely to be presented again, given past frequency.
If a decay rate of 0 is used on the optimized learning form of $B_{i}$, then the equation collapses to $log n_{i}$,
which is a nonlinear transformation of the $B_{i}$ used by \citeauthor{Stanley2013} ($log \frac{p_{i}}{1-p_{i}}$).
Both forms maintain the same rank ordering, so they should produce similar results.
So the $B_{i}$ component used by \citeauthor{Stanley2013} can be roughly thought of as the standard form of $B_{i}$ when the decay rate is 0 (i.e., a pure frequency model).

@@

\begin{figure}[!htbp]
  {%
    \setlength{\fboxsep}{0pt}%
    \setlength{\fboxrule}{1pt}%
    \hfill
    \subfigure[][StackOverflow]{\includegraphics[width=8cm]{HTByTime-520957-crop.pdf}}
    \hfill
    \subfigure[][Twitter]{\includegraphics[width=8cm]{HTByTime-fashionista_com-crop.pdf}}
    \caption{Tagging profile for a single user}
    \label{figPriorHTByTime}
  }%
\end{figure}

The specific value of the hashtag used for both plots is a nominal variable and not important for this analysis.
What is interesting is the large tag space for these users.
There are bands of commonly-used tags for each user, but there are also many instances where new tags are used, and a large percentage of tags used are not in the most common bands.
This helps illustrate how challenging modeling this task can be.
The model's goal is to predict at each instance what tag will be chosen, but to do so the model must predict the correct few tags from a much larger set of tags whose frequency of occurrence is dynamically changing over time.

@@

\begin{figure}[!htbp]
  {%
    \setlength{\fboxsep}{0pt}%
    \setlength{\fboxrule}{1pt}%
    \hfill
    \subfigure[][Standard form]{\includegraphics[width=8cm]{visNormMean-SOQgt500r2-topHashtagPostPriorStd-crop.pdf}}
    \hfill
    \subfigure[][Optimized learning form]{\includegraphics[width=8cm]{visNormMean-SOQgt500r2-topHashtagPostPriorOL2-crop.pdf}}
    \caption{Model performance for a single dataset slice for StackOverflow}
    \label{figPriorSOQSliceDsStd}
  }%
\end{figure}

@@

Aggregate best-fit decay rate values for each model are included in Figure \ref{figPriorDecay}.
Error bars represent the 95\% bootstrapped confidence interval of the median.

\begin{figure}[!htbp]
  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-d--crop.pdf}}}
  \caption{Overall best-fit decay rate for StackOverflow and Twitter}
  \label{figPriorDecay}
\end{figure}

@@

Finally, model accuracy was analyzed across all dataset slices for StackOverflow and Twitter.
Accuracy for both sites is included in Figure \ref{figPriorAcc}. 
Aggregate accuracy was computed by taking the mean of every user's model accuracy across all users in all dataset slices.
All confidence intervals are 95\% bootstrapped for the mean.

\begin{figure}[!htbp]
  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc--crop.pdf}}}
  \caption{Overall model accuracy for StackOverflow and Twitter}
  \label{figPriorAcc}
\end{figure}

@@

@@

The declarative memory models analyzed in this study consist of two main components: past behavior and context.
Past tag use was isolated in the first study to see how well ACT-R's prior component of declarative memory (i.e., base-level learning) fits each user's tag use over time.
For the rest of the studies the past behavior and context declarative memory components were combined as well as looked at in isolation.

@@

In order to best combine these predictors, a logistic regression statistical technique similar to \textcite{Stanley2013} was used to determine the most optimal weights for each term.
Logistic regression is used on these datasets in the following way:
When a user creates a tag for a post, a retrieval request for the model is made.
For each request, the model returns a set of tags (all tags that have been observed for that user in the past) and activations associated with each tag.
An activation value is computed for each model component (e.g., prior tag use, context) for each tag.
These values are recorded, and tag instances that match what the author actually tagged are marked as a 1, while all others are marked with a 0.
This process is repeated across all posts where a user chooses a tag, and across all users in the dataset.
Once all of the results are aggregated, a logistic regression is run where the optimal weights for each model component are found that maximize the model's ability to correctly label tags as 0 or 1 based on total activation.

@@

A simplified version of \citeauthor{Stanley2013}'s method was used for the results in this study.
Instead of subtracting off the mean total activation for the top 5 tags for each post, the process is done in isolation for each model component.
So for the prior component, the activation value used in the logistic regression is the prior activation with the mean of the top 5 prior activations subtracted off of this value.
The same process is used for the context component.
Decoupling the offset term and computing an offset for each model component in isolation means that an iterative approach is no longer necessary,
since the optimal weights of the terms are no longer needed when computing the offset.

@@

\subsubsection{Computing Prior}

@@

The context component for both the random permutation and Bayesian model is based on the same co-occurrence matrix of words and observed tags.
Different size matrices were analyzed, ranging from \num{1000}, \num{10000}, \num{100000}, \num{1000000}, and \num{3000000} posts.
The matrix for \num{3000000} posts was only used for Twitter, since the number of words in a StackOverflow post is much higher than Twitter on average,
and the space and computational time required when using a co-occurrence matrix of \num{3000000} posts for StackOverflow is too high for a current high performance desktop, even with highly optimized representations.
Already \num{835500000} total co-occurring word tag observations are used to build a co-occurrence matrix of \num{1000000} StackOverflow posts.
It is also the case that \num{1000000} posts for StackOverflow is already around $1/7$ of the total posts created on the site,
so it is reasonable to assume that the results would not change greatly for StackOverflow if $3/7$ of the total posts created were used to build the co-occurrence matrix instead,
especially since \num{835500000} co-occurrences are already used with \num{1000000} posts.

@@

Model accuracy relaxed across posts with and without the offset model component was evaluated for both the Bayesian and random permutation model.

@@

The impact of adding the offset to each of the models for StackOverflow and Twitter is shown in Figures \ref{figContextOffsetSO} and \ref{figContextOffsetT}.
The error bars represent the 95\% bootstrapped confidence interval of the mean model accuracy across all runs in dataset.

\begin{figure}[!htbp]
  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-OffsetSO-crop.pdf}}}
  \caption{Impact of adding offset component for StackOverflow}
  \label{figContextOffsetSO}
\end{figure}

\begin{figure}[!htbp]
  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-OffsetT-crop.pdf}}}
  \caption{Impact of adding offset component for Twitter}
  \label{figContextOffsetT}
\end{figure}

In both plots, if the model name includes ``full'' then all model components are included.
When they are enumerated, then at least one of the components is left out.
The random permutation models are represented as ``RP'', Bayesian models as ``Bayes'', and they are plotted alongside each other so that comparisons across the two model types can be done easily.
So, for example, the ``standard prior model relaxed across posts without offset'' means that only the prior model component was used and the offset for that component was not subtracted off when determining activation.
As another example, the ``Bayes only orderless context'' means that only the context component of the Bayesian model was used (along with the offset), and that no order information was used for that contextual component term.

Looking across the plots, including the offset significantly increases model accuracy, for all models, terms, and sites
(\num{.53} to \num{.60} and \num{.22} to \num{.45} for Bayesian and random permutation for StackOverflow, \num{.31} to \num{.29} and \num{.24} to \num{.25} for Bayesian and random permutation for Twitter),
except that the full Bayesian model without an offset was slightly more accurate on Twitter.
Accuracy is improved even when only a single term is used.
Although in this case the weight of the term is irrelevant (only one term, weighting for that term does not matter),
using an offset still improves accuracy since the offset term ensures that only a small number of tags per posts are predicted by the model to be the chosen tags.

@@

It is somewhat disappointing that the offset term is even needed.
If it were the case that model activation for a tag in a specific post could be interpreted as an exact activation value that did not depend on the specific post,
then the model would be able to say when, for example, the third top tag in a post is more likely to be a chosen tag than the second top tag in another post.
That is, the activation values would have meaning more than just rank ordering, and could be used to determine, for example, the number of retrievals to use for a specific post.
If this were the case, then adding the offset term would actually decrease accuracy, since this absolute value of the activation values would be subtracted off.
However, since accuracy improves when adding the offset term for both models, it does not seem that activation values have much meaning across posts (and only within a single post) for these datasets.

A likely and simple explanation for this is that possibly there are posts where a large set of tags have high activation,
but the author is throttling their retrieval threshold for each post such that only a small set of tags are actually used.
This is probably not the retrieval threshold used in the core declarative memory system that determines if a retrieval for the term is successful or not.
It is unlikely that this retrieval threshold can be directly controlled and manipulated.
Rather, it could be a high-level task retrieval threshold that the author has control over, where he/she actively sets it based on competing goals of tagging posts an not overly saturating a post with tags.
Adding the offset term into the model is a way to implement a simple version of this high-level retrieval threshold.

@@

Since the interactions across popular-users datasets are minimal, model behavior for the popular-users subset for Twitter can be interpreted as behavior that will persist across different time periods,
if the same method of choosing popular hashtags is done at a later date.
It may be more general that model behavior also does not depend on the specific method of choosing the popular hashtags, but since only a single method was used for this analysis,
the results from this study cannot by themselves be used to answer this question.

@@

The runs from the Twitter popular-hashtags and StackOverflow randomly-sampled datasets were used for the analysis.
Both the random permutation and Bayesian models were tested.
Activations for context were computed by using the two methods for stop word removal, the entropy weighting method, and no method (i.e., no stop word processing).

The same 571-word Cornell SMART stoplist used by \textcite{Sahlgren2008} was also used for this analysis,
The same entropy weighting method using by \textcite{Stanley2013} was examined,
and the same method of removing the high-frequency words from \citeauthor{Sahlgren2008} was used.

@@

However, a cutoff point must be identified and used in order to remove the high-frequency words.
\textcite{Sahlgren2008} used a cutoff where words that occurred more than \num{15000} times were removed.
Since the StackOverflow and Twitter datasets are much larger than the datasets used by \citeauthor{Sahlgren2008}, the same cutoff could not be used.

@@

The original plot shows the size of the spikes when high-frequency words are not removed.
The entropy plot shows how the size decreases by more than an order of magnitude when the entropy weighting measure is used.
The cutoff for the frequency plot was chosen such that the plot looks similar to the entropy plot, with spikes around the same size.
Note that using a predefined stoplist does not remove some of the remaining high-frequency words in these datasets, which hints that this method may not be ideal when model accuracy is examined.
The cutoff that produced the frequency plots in Figures \ref{figContextCutoffSO} and \ref{figContextCutoffT} is where a word represents more than \num{.04}\% of total occurrences in the dataset.
This may or may not be equal to the value for \textcite{Sahlgren2008}, since that number was expressed in total observations (not as a percentage), and the total number of words in the dataset was not given.

Also, no strong claims are being made that the cutoff identified is the ideal frequency percentage cutoff for all datasets when a frequency cutoff is used.
The method used to identify this cutoff was simply aimed to find a reasonable value where comparisons between a frequency cutoff and entropy weighting techniques would be equitable.

>>>>

@@

These two memory systems were compared across two domains where users generate tags for posts: Twitter and StackOverflow.
Twitter is a microblogging service where users create 140 character tweets and broadcast those messages out to the users who follow their posts.
A user is free to annotate important concepts or words in the tweet by preceding words with hashtags ``\#.''
After the tweet is created the hashtags become hyperlinks, and if clicked on will take a user to a summary feed of all of the tweets across Twitter that have used that hashtag.
So these user-created hashtags on Twitter help connect the tweet to other tweets about the same topic through common hashtag use.
Some recent and often used hashtags for Twitter are \emph{\#photography}, \emph{\#startup}, \emph{\#4change}, \emph{\#android}, and \emph{\#solar}.

@@

Another issue with LSA is that the SVD matrix decomposition technique is computationally very expensive.
As the size of the original word by document matrix increases, the SVD computation becomes more and more time consuming.
If the domains where LSA is applied contained small enough datasets where SVD computation could be computed in a reasonable amount of time, then this would not be much of a concern.
However, with large human-created datasets such as Twitter, StackOverflow, and Wikipedia, the total size of the dataset far exceeds the maximum size that can even be computed by an LSA approach.
\textcite{Budiu2007} used the first six million pages of the Stanford WebBase corpus to compare performance between LSA and PMI on the TOEFL, and the LSA model could not be implemented with this size training dataset.
Instead, they compared the measures on the TOEFL by training on the smaller Touchstone Applied Science Associates (TASA) corpus created from \num{60527} samples of text from high school to early college textbooks.
This is one of the available corpora on the LSA website and is intended to summarize the general reading up to the 1st year of college \parencite{Budiu2007}.

@@

With vector-based systems, order information is included by creating an environment vector ($e_{bind(i,\Phi)}$) that is a function of environment vectors for both the word ($e_{i}$) and location ($\Phi$).
This function has the useful property that the resulting environment vector and the location vector can be inverted to return the original word's environment vector.
Circular convolution and inverse circular convolution (i.e., circular correlation) are used by \textcite{Plate1995} and \textcite{Jones2007} to implement these operations.
This allows for retrieval requests such as ``playing \#$\Phi$,'' ``just read \#$\Phi$,'' and ``vector-based $\Phi$ systems.''

@@

Using decoding, the word-order memory vector for martin is deconvolved to the right ($o_{martin} \circledast \Phi$) which returns the environment vector for luther ($e_{luther}$).
Using resonance, a memory vector for ``martin $\Phi$'' is constructed ($e_{martin} * \Phi$), which correlates highest with the word-order memory vector for luther ($o_{luther}$).
Either way, luther is returned as the most likely term to match the phrase ``martin $\Phi$.''

@@

For example, suppose that the words ``stack overflow'' appear in a sentence, and I want to represent that ``stack'' preceded the word ``overflow.''
The ``stack'' environment vector ($e_{stack}$) is shifted one to the left, producing an environment vector for ``stack'' preceding a word ($e_{stack^{-1}}$).
That environment vector is then added to the memory vector for ``overflow'' ($m_{overflow} = m_{overflow'} + e_{stack^{-1}}$), so that ``overflow'' remembers that ``stack'' has preceded it in a sentence.

@@

The fan effect describes a behavior of human memory where memory items are more difficult to retrieve if they have occurred in a larger number of different contexts.
\textcite{Rutledge2008} used \textcite{Anderson1974} classic experimental paradigm to test the fan effect.
In the task, participants are exposed to a set of person and place pairs (e.g., ``the hippie is in the park'').
They are later asked to recall if a particular person and place pair was presented to them before.
The main manipulation is the ``fan'' of the person and place for each pair.
The fan is defined as the number of different contexts that each item in the pair is presented in.
For example, if items were presented to the participant where the ``hippie'' was both in the ``park'' and in the ``store,'' then ``hippie'' would have a fan of two.
Recall accuracy decrease and time to recall increase as total fan of the person and pair increases.

@@

This research suggests that model fit can improve when tag predictions are customized to the specific user's past tagging history.
It also seems plausible that model accuracy can improve an even greater amount by using the time-dependent form of the base-level component in Table \ref{tabACTRModel},
in order to account for the fact that a user's tagging preferences change over time.

@@

A additional popular-hashtags dataset was collected for Twitter.
The purpose of this dataset is to test the accuracy of each model's predicted most likely hashtag(s) for a tweet.
The dataset contains approximately three million tweets where at least one of the 400 most popular hashtags are used.
A quota of three million tweets was selected because previously research testing Bayesian co-occurrence models on large-scale datasets has used on the order of one million documents
\parencites{Stanley2013,Douglass2010,Budiu2007,Kuo2011}.
Tweets were collected in real time until the tweet quota was reached.

@@

``Stop words'' (e.g., ``the,'' ``a,'' ``or'') are commonly used words that co-occur with all tags (i.e., have little to no discriminating power).
It is typically recommended \parencite{Bird2009} to remove stop words when building the co-occurrence matrix.
Otherwise the matrix may not scale as well since these low-predictor stop words commonly appear in the text.
Also for vector-based systems, memory vectors will become overly saturated with the environment vectors of the commonly-occurring stop words.

@@

For both equations, $i$ represents each chunk (e.g., a word), $j$ is a specific observation for a chunk (e.g., the 2nd time that ``loop'' appears), and $n$ is the number of times each chunk has been observed.
Since the optimized learning form assumes a constant presentation rate for each chunk,
it only has to keep track of the time since the first presentation of a chunk ($L$) and the total number of instances observed for that chunk ($n$). 
It does not have to keep track of the time each specific chunk instance was observed.
Consequently, the optimized learning form of $B_{i}$ is computationally much less expensive than the standard form, and is actually set to the default in the ACT-R Common Lisp implementation.

@@

The base-level ($B_{i}$) component in Table \ref{tabACTRBLLModel} is different than the $B_{i}$ used in \textcite{Stanley2013} (Table \ref{tabModACTRModel}).
\citeauthor{Stanley2013} assumed no time dependence, no decay rate, and that the $B_{i}$ is simply a log-odds computation of the probability that each chunk is likely to be presented again, given past frequency.
If a decay rate of 0 is used on the optimized learning form of $B_{i}$, then the equation collapses to $log \left ( n_{i} \right )$,
which is a nonlinear transformation of the $B_{i}$ used by \citeauthor{Stanley2013} ($log \frac{p_{i}}{1-p_{i}}$).
Both forms maintain the same rank ordering, so they should produce similar results.
So the $B_{i}$ component used by \citeauthor{Stanley2013} can be roughly thought of as the standard form of $B_{i}$ when the decay rate is 0 (i.e., a pure frequency model).

@@

\begin{figure}[!htbp]
  {%
    \setlength{\fboxsep}{0pt}%
    \setlength{\fboxrule}{1pt}%
    \hfill
    \subfigure[][StackOverflow]{\includegraphics[width=8cm]{HTByTime-520957-crop.pdf}}
    \hfill
    \subfigure[][Twitter]{\includegraphics[width=8cm]{HTByTime-fashionista_com-crop.pdf}}
    \caption{
      Tagging profile for a single user.
      The x-axis represents the specific hashtag a user chooses for a post and time progresses vertically.
      A point in the plot represents when a user chooses a specific hashtag for a specific post.
      A user's chosen hashtag for their oldest post that contains a hashtag is the lowest point on the plot.
    }
    \label{figPriorHTByTime}
  }%
\end{figure}

The specific value of the hashtag used for both plots is a nominal variable and not important for this analysis.
What is interesting is the large tag space for these users.
There are bands of commonly-used tags for each user, but there are also many instances where new tags are used, and a large percentage of tags used are not in the most common bands.
For example, the StackOverflow user in Figure \ref{figPriorHTByTime} has a pronounced band where the same hashtag is chosen repeatedly over time. 
However there are many instances of other tags being used in parallel to the tag that is producing the band (that is repeatedly being used). 
The Twitter user has a few clusters where the same hashtag is chosen repeatedly but only for a brief period of time.
As time progresses, a band of choosing another hashtag is observed, and then the user switches to a new hashtag after that.
More recent posts for the Twitter user are more varied, and there is no longer a single hashtag that the user chooses repeatedly.
This helps illustrate how challenging modeling this task can be.
The model's goal is to predict at each instance what tag will be chosen, but to do so the model must predict the correct few tags from a much larger set of tags whose frequency of occurrence is changing over time.

@@

\begin{figure}[!htbp]
  {%
    \setlength{\fboxsep}{0pt}%
    \setlength{\fboxrule}{1pt}%
    \hfill
    \subfigure[][Standard form]{\includegraphics[width=8cm]{visNormMean-SOQgt500r2-topHashtagPostPriorStd-crop.pdf}}
    \hfill
    \subfigure[][Optimized learning form]{\includegraphics[width=8cm]{visNormMean-SOQgt500r2-topHashtagPostPriorOL2-crop.pdf}}
    \caption{
    Model performance for a single dataset slice for StackOverflow.
  }
    \label{figPriorSOQSliceDsStd}
  }%
\end{figure}

@@

Aggregate best-fit decay rate values for each model are included in Figure \ref{figPriorDecay}.

\begin{figure}[!htbp]
  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-d--crop.pdf}}}
  \caption{
  Overall best-fit decay rate for StackOverflow and Twitter.
  Results are shown for the two Twitter and StackOverflow (SO) popular-users datasets.
  Error bars represent the 95\% bootstrapped confidence interval for the median.
}
  \label{figPriorDecay}
\end{figure}

@@

Finally, model accuracy was analyzed across all dataset slices for StackOverflow and Twitter.
Accuracy for both sites is included in Figure \ref{figPriorAcc}. 
Aggregate accuracy was computed by taking the mean of every user's model accuracy across all users in all dataset slices.

\begin{figure}[!htbp]
  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc--crop.pdf}}}
  \caption{
    Overall model accuracy for StackOverflow and Twitter at optimal decay-rate values.
    Best-fit decay-rate values for the standard and optimized learning model are \num{.8} and \num{.43}.
    All error bars are the 95\% bootstrapped confidence interval for the mean. 
  }
  \label{figPriorAcc}
\end{figure}

@@

The decay rate parameter in $B_{i}$ determines how fast a presentation of a chunk in the past decays over time.
Higher values of $d$ mean that activation decays more quickly, which means that the chunks used more recently are the most active and the total frequency count of a chunk matters less.
Looking back at Figure \ref{figPriorDecay}, the optimal decay rate parameter when the standard form of $B_{i}$ is used is higher for Twitter than StackOverflow.
It makes sense that the lifetime of a specific hashtag on Twitter is shorter overall than the lifetime of a tag (e.g., \emph{C#}) on StackOverflow.
This means that the decay rate for Twitter should be higher than StackOverflow, which is what is observed.

@@

The declarative memory models analyzed for this research consist of two main components: past behavior and context.
Past tag use was isolated in the first study to see how well ACT-R's prior component of declarative memory (i.e., base-level learning) fits each user's tag use over time.
For the rest of the studies the past behavior and context declarative memory components were combined as well as looked at in isolation.

@@

In order to best combine these predictors, a logistic regression statistical technique similar to \textcite{Stanley2013} was used to determine the most optimal weights for each term.
Each term in the full model (see the total activation equation Table \ref{tabModACTRModel}) is weighted so that model accuracy is maximized.  
\textcite{Stanley2013} found that the optimal weights for the StackOverflow model for the prior term, title context term, and body context term were \num{.62}, \num{.93}, and \num{1.75} respectively.

The optimal weights for each model component are computed in the following way:
When a user creates a tag for a post, a retrieval request for the model is made.
For each request, the model returns a set of tags (all tags that have been observed for that user in the past) and activations associated with each tag.
An activation value is computed for each model component (e.g., prior tag use, context) for each tag.
These values are recorded, and tag instances that match what the author actually tagged are marked as a 1, while all others are marked with a 0.
This process is repeated across all posts where a user chooses a tag, and across all users in the dataset.
Once all of the results are aggregated, a logistic regression is run where the optimal weights for each model component are found that maximize the model's ability to correctly label tags as 0 or 1 based on total activation.

@@

A simplified version of \citeauthor{Stanley2013}'s method was used for the results for this research.
Instead of subtracting off the mean total activation for the top 5 tags for each post, the process is done in isolation for each model component.
So for the prior component, the activation value used in the logistic regression is the prior activation with the mean of the top 5 prior activations subtracted off of this value.
The same process is used for the context component.
Decoupling the offset term and computing an offset for each model component in isolation means that an iterative approach is no longer necessary,
since the optimal weights of the terms are no longer needed when computing the offset.

@@

\subsubsection{Computing Priors}

@@

The context component for both the random permutation and Bayesian model is based on the same co-occurrence matrix of words and observed tags.
Different size matrices were analyzed, ranging from \num{1000}, \num{10000}, \num{100000}, \num{1000000}, and \num{3000000} posts.
The matrix for \num{3000000} posts was only used for Twitter, since the number of words in a StackOverflow post is much higher than Twitter on average,
and the space and computational time required when using a co-occurrence matrix of \num{3000000} posts for StackOverflow is too high for a current high performance desktop, even with highly optimized representations.
Already \num{835500000} total co-occurring word tag observations were used to build a co-occurrence matrix of \num{1000000} StackOverflow posts.
It was also the case that \num{1000000} posts for StackOverflow was already around $1/7$ of the total posts created on the site,
so it is reasonable to assume that the results would not change greatly for StackOverflow if $3/7$ of the total posts created were used to build the co-occurrence matrix instead,
especially since \num{835500000} co-occurrences were already used with \num{1000000} posts.

@@

The offset component was added to both models, and model accuracy was examined before and after the addition.

@@

The impact of adding the offset to each of the models for StackOverflow and Twitter is shown in Figures \ref{figContextOffsetSO} and \ref{figContextOffsetT}.

\begin{figure}[!htbp]
  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-OffsetSO-crop.pdf}}}
  \caption{
    Impact of adding offset component for StackOverflow.
    Error bars represent the 95\% bootstrapped confidence interval for the mean model accuracy across all runs in dataset.
    The random permutation models are represented as ``RP'' and Bayesian models as ``Bayes.''
}
  \label{figContextOffsetSO}
\end{figure}

\begin{figure}[!htbp]
  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-OffsetT-crop.pdf}}}
  \caption{
    Impact of adding offset component for Twitter.
    All error bars are the 95\% bootstrapped confidence interval for the mean.
  }
  \label{figContextOffsetT}
\end{figure}

In both plots, if the model name includes ``full'' then all model components are included.
When they are enumerated, then at least one of the components is left out.
The random permutation and Bayesian models are plotted alongside each other so that comparisons across the two model types can be done easily.
So, for example, the ``standard prior model relaxed across posts without offset'' means that only the prior model component was used and the offset for that component was not subtracted off when determining activation.
As another example, the ``Bayes only orderless context'' means that only the context component of the Bayesian model was used (along with the offset), and that no order information was used for that contextual component term.

Looking across the plots, including the offset significantly increases model accuracy, for all models, terms, and sites
(\num{.53} to \num{.60} and \num{.22} to \num{.45} for Bayesian and random permutation for StackOverflow, \num{.31} to \num{.29} and \num{.24} to \num{.25} for Bayesian and random permutation for Twitter),
except that the full Bayesian model without an offset was slightly more accurate on Twitter.
Accuracy is improved even when only a single term is used.
Although in this case the term is not weighted,
but using an offset still improves accuracy since the offset term ensures that only a small number of tags per posts are predicted by the model to be the chosen tags.

@@

It is somewhat disappointing that the offset term is even needed.
If it were the case that model activation for a tag in a specific post could be interpreted as an exact activation value that did not depend on the specific post,
then the model would be able to say when, for example, the third top tag in a post is more likely to be a chosen tag than the second top tag in another post.
That is, the activation values would have meaning more than just rank ordering, and could be used to determine, for example, the number of tags to use for a specific post.
If this were the case, then adding the offset term would actually decrease accuracy, since this absolute value of the activation values would be removed.
However, since accuracy improves when adding the offset term for both models, it does not seem that absolute activation values have much meaning across posts for these datasets,
and that the relative activation values for each tag within a post is what is important.

One simple and likely explanation is that there are posts where a large set of tags have high activation,
but the author is choosing to use only a fixed small set of the top tags for each post, regardless of the activation values of the top tags.
The author determines the number of tags to use for a post based on competing goals of tagging posts an not overly saturating a post with tags.
For example, the author may be writing a question on StackOverflow about \emph{javascript}, and a large set of related tags are retrieved from declarative memory that are all above the retrieval threshold
(e.g., \emph{javascript}, \emph{dom}, \emph{svg}, \emph{html}, \emph{jquery}, \emph{css}),
but the author only chooses to use the tags \emph{javascript}, \emph{dom}, and \emph{svg}, since he has decided to tag this post with three tags, and those are the three tags with the highest activation.
Adding the offset term into the model is a way to implement a simple version of this strategy where the author is actively choosing to use only a small set of tags for each post.

@@

Since the interactions across popular-users datasets are minimal, model behavior for the popular-users subset for Twitter can be interpreted as behavior that will persist across different time periods,
if the same method of choosing popular hashtags is conducted at a later date.
It may also be the case that that model behavior does not depend on the specific method of choosing the popular hashtags, but since only a single method was used for this analysis,
the results from this study cannot by themselves be used to answer this question.

@@

The runs from the Twitter popular-hashtags and StackOverflow randomly-sampled datasets were used for the analysis.
Both the random permutation and Bayesian models were tested.
Activations for context were computed by using five different methods to handle stop words:
The two removal methods used in \textcite{Sahlgren2008} (i.e., removal based on frequency, removal based on the pre-determined 571-word Cornell SMART stoplist),
the entropy-weighting method used in \textcite{Stanley2013},
combining the entropy-weighting and frequency-filtering methods,
and no removal or weighting of stop words.

@@

A cutoff point must be identified and used in order to remove the high-frequency words.
\textcite{Sahlgren2008} used a cutoff where words that occurred more than \num{15000} times were removed.
Since the StackOverflow and Twitter datasets are much larger than the datasets used by \citeauthor{Sahlgren2008}, the same cutoff could not be used.

@@

The original plot shows the size of the spikes when high-frequency words are not removed.
The entropy plot shows how the size decreases by more than an order of magnitude after the entropy weighting measure is used (compare the difference between the original subplot and the entropy subplot).
The cutoff for the frequency plot was chosen such that the plot looks similar to the entropy plot, with spikes around the same size.
The cutoff that produced the frequency plots in Figures \ref{figContextCutoffSO} and \ref{figContextCutoffT} is where a word represents more than \num{.04}\% of total occurrences in the dataset.
This may or may not be equal to the value for \textcite{Sahlgren2008}, since that number was expressed in total observations (not as a percentage), and the total number of words in the dataset was not given.

Note that using a predefined stoplist does not remove some of the remaining high-frequency words in these datasets, which hints that this method may not be ideal when model accuracy is examined.
Also, no strong claims are being made that the cutoff identified is the ideal frequency percentage cutoff for all datasets when a frequency cutoff is used.
The method used to identify this cutoff was simply intended to find a reasonable value where comparisons between a frequency cutoff and entropy weighting techniques would be equitable.
